---
title: "Summarize Variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarize Variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rubix)
library(cave)
data("mtcars")
```

# Introduction  
`Summarize_` wrappers to the dplyr functions of the same name were developed to quickly QA the depth and breadth of data in a given dataframe. The most comprehensive function in this family of functions are the summarize_variables function, where counts are returned as well as a string of column names that can be easily converted back into a vector using the `string_to_vector()` function in the cave R package. 

```{r summarize_variables function}
rubix::summarize_variables(mtcars)
```

Summary functions include:  
1. Total and distinct counts  
2. Counts for NA (ie. <NA>), NA strings (ie "NA"), and blank values that when combined, can provide an overview of missingness in the dataframe  
3. Unique values found within the particular variable in both a pipe-separated string as well as a R expression that can be evaluated back into a vector using the `string_to_vector()` function in the cave R package.  

The scope of the summary may also be limited by variable: 
```{r}
rubix::summarize_variables(mtcars, mpg, cyl)
```


## Summarize Numeric Variables  
Additional metrics can be derived from variables that contain numeric data. The `summarize_numeric_var()` function either takes the provided variables or selects for variables of numeric, integer, or double classes and calculates summary statistics with both na.rm = FALSE (all _NA suffixed outputs) and na.rm = TRUE parameters. 
```{r}
rubix::summar
```




