---
title: "Summarize Variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarize Variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rubix)
data("mtcars")
```

# Introduction  
`Summarize_` functions rapidly survey the depth and breadth of a dataframe. The general use cases for this function family is to analyze data for variability from multiple angles to bucket values into broader categories. 

```{r summarize_variables function}
summarize_variables(mtcars,
                    incl_num_calc = FALSE)
```

Summary functions include:  
1. Total and distinct counts  
2. Counts for NA (ie. <NA>), NA strings (ie "NA"), and blank values that when combined, can provide an overview of missingness in the dataframe  
3. Unique values found within the particular variable in both a pipe-separated string .  


## Summarize Numeric Variables  
Additional metrics can be derived from variables that contain numeric data. The `summarize_variables()` function either takes variables as arguments or selects for variables of numeric, integer, or double R classes and calculates summary statistics with both na.rm = FALSE (all _NA suffixed outputs) and na.rm = TRUE parameters.  

```{r summarize_numeric_var}
summarize_variables(data = mtcars,
                    incl_num_calc = TRUE)
```

# Summarizing Variable Values  
The `value_count()` function returns all counts for the unique values for each variable. 

```{r}
value_count(data = mtcars)
```


